Determining the existance and direction of causality in the face of measurement error

### 2016-12-13 docs/manuscript.tex
upated

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Determining the existance and direction of causality in the face of measurement error},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\setlength{\droptitle}{-2em}
  \title{Determining the existance and direction of causality in the face of
measurement error}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{09 May 2016}

\begin{document}

\maketitle

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \(H_0: cov(g, x) = 0; H_1: cov(g, x) \neq 0\); \emph{the SNP
  associates with the exposure}
\item
  \(H_0: cov(g, y) = 0; H_1: cov(g, y) \neq 0\); \emph{the SNP
  associates with the outcome}
\item
  \(H_0: cov(x, y) = 0; H_1: cov(x, y) \neq 0\); \emph{the exposure
  associates with the outcome}
\item
  \(H_0: cov(g, y - \hat{y}) \neq 0; cov(g, y - \hat{y}) = 0\);
  \emph{the SNP is independent of the outcome when the outcome is
  adjusted for the exposure}
\end{enumerate}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  If \(p_{CIT,correct} < \alpha\) and \(p_{CIT,incorrect} > \alpha\)
  then the correct model is accepted
\item
  If \(p_{CIT,correct} > \alpha\) and \(p_{CIT,incorrect} < \alpha\)
  then the incorrect model is accepted
\item
  If \(p_{CIT,correct} > \alpha\) and \(p_{CIT,incorrect} > \alpha\)
  then an alternative causal model is accepted
\item
  If \(p_{CIT,correct} < \alpha\) and \(p_{CIT,incorrect} < \alpha\)
  then no inference is made
\end{itemize}

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  If \(p_{Steiger} < \alpha\) and \(p_{MR} < \alpha\) and \(Z > 0\) then
  a causal association for the correct model is accepted
\item
  If \(p_{Steiger} < \alpha\) and \(p_{MR} < \alpha\) and \(Z < 0\) then
  a causal association for the incorrect model is accepted
\item
  Otherwise no inference is made
\end{itemize}

@@

We used the Steiger test to test if DNA methylation levels are more
likely to causally influence gene expression levels or vice versa. To do
this we obtained a list of 458 gene expression - DNA methylation
associations as reported in Shakhbazov et al\textsuperscript{18}. These
were filtered to be located on the same chromosome, have robost
correlations after correcting for multiple testing, and to share a SNP
that had a robust cis-acting effect on both the DNA methylation probe
and the gene expression probe. Using the reported summary data we used
the Steiger test of two independent correlations to identify which
direction of causality was most likely for each association. We found
that the causal direction commonly goes in both directions (Table 1) but
methylation levels were more likely to influence expression levels than
vice versa (probability of expression causing methylation = 0.0024038
(95\% CI 0.3446222 - 0.4667222) amongst 262 associations with
\(p_{Steiger} < 0.05\)).

@@

We also observed that for associations where methylation caused gene
expression the causal effect was more likely to be negative than for the
associations where gene expression caused methylation (OR = 0.505822
(95\% CI 0.2813174 - 0.89394)), suggesting that reducing gene expression
levels at a controlling CpG typically leads to increased gene expression
levels, consistent with expectation\textsuperscript{2}.

@@

\begin{longtable}[c]{@{}lrrrr@{}}
\toprule
Causal direction & Count & Count (\$p\_\{Steiger\} \textless{} 0.05) &
P(+ve effect) & \(cor(\rho_{MR}, \rho_{P})\)\tabularnewline
\midrule
\endhead
Expression causes Methylation & 246 & 156 & 0.3902439 &
0.5629453\tabularnewline
Methylation causes Expression & 212 & 106 & 0.2877358 &
0.4384660\tabularnewline
\bottomrule
\end{longtable}

@@

@@

1.Kettunen, J. \emph{et al.} Genome-wide association study identifies
multiple loci influencing human serum metabolite levels. \emph{Nature
genetics} \textbf{44,} 269--76 (2012).

2.Bird, A. DNA methylation patterns and epigenetic memory. \emph{Genes
\& development} \textbf{16,} 6--21 (2002).

3.{Davey Smith}, G. \& Ebrahim, S. Mendelian randomization: prospects,
potentials, and limitations. \emph{International journal of
epidemiology} \textbf{33,} 30--42 (2004).

4.Baron, R. M. \& Kenny, D. A. The moderator-mediator variable
distinction in social psychological research: conceptual, strategic, and
statistical considerations. \emph{Journal of personality and social
psychology} \textbf{51,} 1173--82 (1986).

5.Aten, J. E., Fuller, T. F., Lusis, A. J. \& Horvath, S. Using genetic
markers to orient the edges in quantitative trait networks: the NEO
software. \emph{BMC systems biology} \textbf{2,} 34 (2008).

6.Fewell, Z., {Davey Smith}, G. \& Sterne, J. A. C. The impact of
residual and unmeasured confounding in epidemiologic studies: a
simulation study. \emph{American journal of epidemiology} \textbf{166,}
646--55 (2007).

7.Schadt, E. E. \emph{et al.} An integrative genomics approach to infer
causal associations between gene expression and disease. \emph{Nature
Genetics} \textbf{37,} 710--717 (2005).

8.Millstein, J., Zhang, B., Zhu, J. \& Schadt, E. E. Disentangling
molecular relationships with a causal inference test. \emph{BMC
genetics} \textbf{10,} 23 (2009).

9.Millstein, J. \emph{cit: Causal Inference Test. R package version
1.9}. (2016). at
\textless{}\url{http://cran.r-project.org/package=cit}\textgreater{}

10.Waszak, S. M. \emph{et al.} Variation and genetic control of
chromatin architecture in humans. \emph{Cell} \textbf{162,} 1039--1050
(2015).

11.Koestler, D. C. \emph{et al.} Integrative genomic analysis identifies
epigenetic marks that mediate genetic risk for epithelial ovarian
cancer. \emph{BMC medical genomics} \textbf{7,} 8 (2014).

12.Liu, Y. \emph{et al.} Epigenome-wide association data implicate DNA
methylation as an intermediary of genetic risk in rheumatoid arthritis.
\emph{Nature biotechnology} \textbf{31,} 142--7 (2013).

13.Yuan, W. \emph{et al.} An integrated epigenomic analysis for type 2
diabetes susceptibility loci in monozygotic twins. \emph{Nature
communications} \textbf{5,} 5719 (2014).

14.Tang, Y. \emph{et al.} Genotype-based treatment of type 2 diabetes
with an \(\alpha\)2A-adrenergic receptor antagonist. \emph{Science
translational medicine} \textbf{6,} 257ra139 (2014).

15.Hong, X. \emph{et al.} Genome-wide association study identifies
peanut allergy-specific loci and evidence of epigenetic mediation in US
children. \emph{Nature communications} \textbf{6,} 6304 (2015).

16.Claussnitzer, M. \emph{et al.} FTO Obesity Variant Circuitry and
Adipocyte Browning in Humans. \emph{The New England journal of medicine}
\textbf{373,} 895--907 (2015).

17.Timpson, N. J. \emph{et al.} C-reactive protein levels and body mass
index: elucidating direction of causation through reciprocal Mendelian
randomization. \emph{International journal of obesity (2005)}
\textbf{35,} 300--8 (2011).

18.Shakhbazov, K. \emph{et al.} Shared genetic control of expression and
methylation in peripheral blood. \emph{BMC genomics} \textbf{17,} 278
(2016).

19.Pierce, B. L. \& Burgess, S. Efficient design for Mendelian
randomization studies: subsample and 2-sample instrumental variable
estimators. \emph{American journal of epidemiology} \textbf{178,}
1177--84 (2013).

20.Ashenfelter, O. \& Krueger, A. B. Estimates of the Economic Return to
Schooling from a New Sample of Twins. \emph{The American Economic
Review} \textbf{84,} 1157--1173 (1994).

21.Cessie, S. le, Debeij, J., Rosendaal, F. R., Cannegieter, S. C. \&
Vandenbroucke, J. P. Quantification of bias in direct effects estimates
due to different types of measurement error in the mediator.
\emph{Epidemiology (Cambridge, Mass.)} \textbf{23,} 551--60 (2012).

22.Nagarajan, R. \& Scutari, M. Impact of noise on molecular network
inference. \emph{PloS one} \textbf{8,} e80735 (2013).

23.Shpitser, I., VanderWeele, T. \& Robins, J. On the validity of
covariate adjustment for estimating causal effects. \emph{Proceedings of
the Twenty Sixth Conference on Uncertainty in Artificial Intelligence
(UAI-10)} 527--536 (2010).

24.Blakely, T., McKenzie, S. \& Carter, K. Misclassification of the
mediator matters when estimating indirect effects. \emph{Journal of
epidemiology and community health} \textbf{67,} 458--66 (2013).

25.Wang, L. \& Michoel, T. Detection of regulator genes and eQTLs in
gene networks. \emph{arXiv} \textbf{arXiv:1512,} (2015).

26.Lagani, V., Triantafillou, S., Ball, G., Tegner, J. \& Tsamardinos,
I. in \emph{Uncertainty in biology: A computational modeling approach}
47 (Springer, 2015). at
\textless{}\url{https://books.google.com/books?id=8SLUCgAAQBAJ\{/\&\}pgis=1}\textgreater{}

27.Houle, D., P{é}labon, C., Wagner, G. \& Hansen, T. Measurement and
meaning in biology. \emph{The Quarterly Review of Biology} \textbf{86,}
3--34 (2011).

28.Hern{á}n, M. a \& Cole, S. R. Invited Commentary: Causal diagrams and
measurement bias. \emph{American journal of epidemiology} \textbf{170,}
959--62; discussion 963--4 (2009).

29.Harper, K. N., Peters, B. a \& Gamble, M. V. Batch effects and
pathway analysis: two potential perils in cancer studies involving DNA
methylation array analysis. \emph{Cancer epidemiology, biomarkers \&
prevention : a publication of the American Association for Cancer
Research, cosponsored by the American Society of Preventive Oncology}
\textbf{22,} 1052--60 (2013).

30.Chen, Y.-a. \emph{et al.} Discovery of cross-reactive probes and
polymorphic CpGs in the Illumina Infinium HumanMethylation450
microarray. \emph{Epigenetics : official journal of the DNA Methylation
Society} \textbf{8,} 203--9 (2013).

31.Houseman, E. A. \emph{et al.} DNA methylation arrays as surrogate
measures of cell mixture distribution. \emph{BMC bioinformatics}
\textbf{13,} 86 (2012).

32.Ahima, R. S. \& Lazar, M. A. Physiology. The health risk of
obesity--better metrics imperative. \emph{Science (New York, N.Y.)}
\textbf{341,} 856--8 (2013).

33.R Core Team. \emph{R: A Language and Environment for Statistical
Computing}. (R Foundation for Statistical Computing, 2015). at
\textless{}\url{https://www.r-project.org/}\textgreater{}

34.Henningsen, A. \& Hamann, J. D. systemfit : A Package for Estimating
Systems of Simultaneous Equations in R. \emph{Journal of Statistical
Software} \textbf{23,} 1--40 (2007).

35.Steiger, J. H. Tests for comparing elements of a correlation matrix.
\emph{Psychological Bulletin} \textbf{87,} 245--251 (1980).

36.Revelle, W. \emph{psych: Procedures for Psychological, Psychometric,
and Personality Research}. (Northwestern University, 2015). at
\textless{}\url{http://cran.r-project.org/package=psych}\textgreater{}

37.Relton, C. L. \& {Davey Smith}, G. Two-step epigenetic Mendelian
randomization: a strategy for establishing the causal role of epigenetic
processes in pathways to disease. \emph{International journal of
epidemiology} \textbf{41,} 161--76 (2012).

38.Burgess, S., Freitag, D. F., Khan, H., Gorman, D. N. \& Thompson, S.
G. Using multivariable Mendelian randomization to disentangle the causal
effects of lipid fractions. \emph{PloS one} \textbf{9,} e108891 (2014).

39.Burgess, S., Daniel, R. M., Butterworth, A. S. \& Thompson, S. G.
Network Mendelian randomization: using genetic variants as instrumental
variables to investigate mediation in causal pathways.
\emph{International journal of epidemiology} \textbf{44,} 484--95
(2015).

40.Varbo, A. \emph{et al.} Remnant cholesterol, low-density lipoprotein
cholesterol, and blood pressure as mediators from obesity to ischemic
heart disease. \emph{Circulation research} \textbf{116,} 665--73 (2015).

41.Bowden, J., {Davey Smith}, G. \& Burgess, S. Mendelian randomization
with invalid instruments: effect estimation and bias detection through
Egger regression. \emph{International Journal of Epidemiology}
\textbf{In press,} (2015).

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Determining the existance and direction of causality in the face of measurement error},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{Determining the existance and direction of causality in the face of
measurement error}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{13 December 2016}

\begin{document}
\maketitle

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \(H_0: cov(g, x) = 0; H_1: cov(g, x) \neq 0\); \emph{the SNP
  associates with the exposure}
\item
  \(H_0: cov(g, y) = 0; H_1: cov(g, y) \neq 0\); \emph{the SNP
  associates with the outcome}
\item
  \(H_0: cov(x, y) = 0; H_1: cov(x, y) \neq 0\); \emph{the exposure
  associates with the outcome}
\item
  \(H_0: cov(g, y - \hat{y}) \neq 0; cov(g, y - \hat{y}) = 0\);
  \emph{the SNP is independent of the outcome when the outcome is
  adjusted for the exposure}
\end{enumerate}

@@

\begin{itemize}
\tightlist
\item
  If \(p_{CIT,correct} < \alpha\) and \(p_{CIT,incorrect} > \alpha\)
  then the correct model is accepted
\item
  If \(p_{CIT,correct} > \alpha\) and \(p_{CIT,incorrect} < \alpha\)
  then the incorrect model is accepted
\item
  If \(p_{CIT,correct} > \alpha\) and \(p_{CIT,incorrect} > \alpha\)
  then an alternative causal model is accepted
\item
  If \(p_{CIT,correct} < \alpha\) and \(p_{CIT,incorrect} < \alpha\)
  then no inference is made
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  If \(p_{Steiger} < \alpha\) and \(p_{MR} < \alpha\) and \(Z > 0\) then
  a causal association for the correct model is accepted
\item
  If \(p_{Steiger} < \alpha\) and \(p_{MR} < \alpha\) and \(Z < 0\) then
  a causal association for the incorrect model is accepted
\item
  Otherwise no inference is made
\end{itemize}

@@

We used the Steiger test to test if DNA methylation levels are more
likely to causally influence gene expression levels or vice versa. To do
this we obtained a list of 458 gene expression - DNA methylation
associations as reported in Shakhbazov et al\textsuperscript{18}. These
were filtered to be located on the same chromosome, have robost
correlations after correcting for multiple testing, and to share a SNP
that had a robust cis-acting effect on both the DNA methylation probe
and the gene expression probe. Using the reported summary data we used
the Steiger test of two independent correlations to identify which
direction of causality was most likely for each association. We found
that the causal direction commonly goes in both directions (Table 1) but
methylation levels were more likely to influence expression levels than
vice versa (probability of expression causing methylation = 0.0019317
(95\% CI 0.3455756 - 0.4653544) amongst 272 associations with
\(p_{Steiger} < 0.05\)).

@@

We also observed that for associations where methylation caused gene
expression the causal effect was more likely to be negative than for the
associations where gene expression caused methylation (OR = 0.4615363
(95\% CI 0.259998 - 0.8052745)), suggesting that reducing gene
expression levels at a controlling CpG typically leads to increased gene
expression levels, consistent with expectation\textsuperscript{2}.

@@

\begin{longtable}[]{@{}lrrrr@{}}
\toprule
Causal direction & Count & \(p_{Steiger} < 0.05\) & P(+ve effect) &
\(cor(\rho_{MR}, \rho_{P})\)\tabularnewline
\midrule
\endhead
Expression causes Methylation & 246 & 162 & 0.3902439 &
0.5629453\tabularnewline
Methylation causes Expression & 212 & 110 & 0.2877358 &
0.4384660\tabularnewline
\bottomrule
\end{longtable}

@@

\subsection{Appendix 3}\label{appendix-3}

Sensitivity analysis for the Steiger test involves evaluating the
probability that the inferred causal direction is incorrect due to
differences in measurement error between the hypothesised exposure,
\(x\), and hypothesised outcome, \(y\), variables. The objective is to
calculate what the true \(\rho_{g, x}\) and \(\rho_{g, y}\) values would
be given empirical estimates of \(\rho_{g, x_O}\) and \(\rho_{g, y_O}\)
across all possible values of measurement error.

\[
\rho_{g,x} = \frac{\rho_{x,x_O}}{\rho_{g,x_O}}
\]

\[
\rho_{g,y} = \frac{\rho_{y,y_O}}{\rho_{g,y_O}}
\]

If \(\rho_{g,x} > \rho_{g,y}\) then we infer that the causal direction
is from exposure to outcome. Assume that all measurement error values
for \(x\) and \(y\) are equally probable, the surface of possible values
of this inequality is
\(z = \frac{\rho_{y,y_O}}{\rho_{g,y_O}} - \frac{\rho_{x,x_O}}{\rho_{g,x_O}}\),
and the objective is to calculate \(P(z > 0)\), which corresponds to the
proportion of measurement error values that would lead to the Steiger
test being incorrect.

The surface area of \(z\) across the domain
\(\rho_{g,y_O} \leq \rho_{y,y_O} \leq 1; \rho_{g,x_O} \leq \rho_{x,x_O} \leq 1\)
is obtained from

\[
\begin{aligned}
S_z & = \int^1_{\rho_{g,x_o}} \int^1_{\rho_{g,y_o}} \sqrt{  \frac{\partial}{\partial \rho_{x,x_o}} \left (\frac{\rho_{g,x_O}}{\rho_{x,x_O}} \right )^2 +  \left (\frac{\rho_{g,y_O}}{\rho_{y,y_O}} \right )^2   +1 }\, dydx \\
& = \int^1_{\rho_{g,x_o}} \int^1_{\rho_{g,y_o}} \sqrt{ \frac{\rho_{g,x_O}^2}{\rho_{x,x_O}^4} + \frac{\rho_{g,y_O}^2}{\rho_{y,y_O}^4} + 1}\, dydx
\end{aligned}
\]

and the proportion of the surface for which \(z > 0\) is obtained from

\[
\begin{aligned}
S_{z>0} = \int^1_{\rho_{g,x_o}} \int^{R}_{\rho_{g,y_o}} \sqrt{ \frac{\rho_{g,x_O}^2}{\rho_{x,x_O}^4} + \frac{\rho_{g,y_O}^2}{\rho_{y,y_O}^4} + 1}\, dydx
\end{aligned}
\]

where \(R\) is the function for which
\(0 = \frac{\rho_{y,y_O}}{\rho_{g,y_O}} - \frac{\rho_{x,x_O}}{\rho_{g,x_O}}\),
i.e.

\[
R = \frac{\rho_{g,y_o}\rho_{g,x}}{\rho_{g,x_o}}
\]

Finally, the probability that the inference of \(x \rightarrow y\)
obtained from the Steiger test is incorrect due to measurement error is
obtained from

\[
P(M) = \frac{S_{z>0}}{S_z}
\]

\(P(M)\) can range from 0 to 1, where a small value indicates that the
Steiger test is relatively insensitive to measurement error. This value
is estimated through simulation.

@@

\hypertarget{refs}{}
\hypertarget{ref-Kettunen2012}{}
1. Kettunen, J. \emph{et al.} Genome-wide association study identifies
multiple loci influencing human serum metabolite levels. \emph{Nature
genetics} \textbf{44,} 269--76 (2012).

\hypertarget{ref-Bird2002}{}
2. Bird, A. DNA methylation patterns and epigenetic memory. \emph{Genes
\& development} \textbf{16,} 6--21 (2002).

\hypertarget{ref-DaveySmith2004}{}
3. Davey Smith, G. \& Ebrahim, S. Mendelian randomization: prospects,
potentials, and limitations. \emph{International journal of
epidemiology} \textbf{33,} 30--42 (2004).

\hypertarget{ref-Baron1986}{}
4. Baron, R. M. \& Kenny, D. A. The moderator-mediator variable
distinction in social psychological research: conceptual, strategic, and
statistical considerations. \emph{Journal of personality and social
psychology} \textbf{51,} 1173--82 (1986).

\hypertarget{ref-Aten2008}{}
5. Aten, J. E., Fuller, T. F., Lusis, A. J. \& Horvath, S. Using genetic
markers to orient the edges in quantitative trait networks: the NEO
software. \emph{BMC systems biology} \textbf{2,} 34 (2008).

\hypertarget{ref-Fewell2007}{}
6. Fewell, Z., Davey Smith, G. \& Sterne, J. A. C. The impact of
residual and unmeasured confounding in epidemiologic studies: a
simulation study. \emph{American journal of epidemiology} \textbf{166,}
646--55 (2007).

\hypertarget{ref-Schadt2005}{}
7. Schadt, E. E. \emph{et al.} An integrative genomics approach to infer
causal associations between gene expression and disease. \emph{Nature
Genetics} \textbf{37,} 710--717 (2005).

\hypertarget{ref-Millstein2009}{}
8. Millstein, J., Zhang, B., Zhu, J. \& Schadt, E. E. Disentangling
molecular relationships with a causal inference test. \emph{BMC
genetics} \textbf{10,} 23 (2009).

\hypertarget{ref-Millstein2016}{}
9. Millstein, J. \emph{cit: Causal Inference Test. R package version
1.9}. (2016). at
\textless{}\url{http://cran.r-project.org/package=cit}\textgreater{}

\hypertarget{ref-Waszak2015}{}
10. Waszak, S. M. \emph{et al.} Variation and genetic control of
chromatin architecture in humans. \emph{Cell} \textbf{162,} 1039--1050
(2015).

\hypertarget{ref-Koestler2014}{}
11. Koestler, D. C. \emph{et al.} Integrative genomic analysis
identifies epigenetic marks that mediate genetic risk for epithelial
ovarian cancer. \emph{BMC medical genomics} \textbf{7,} 8 (2014).

\hypertarget{ref-Liu2013}{}
12. Liu, Y. \emph{et al.} Epigenome-wide association data implicate DNA
methylation as an intermediary of genetic risk in rheumatoid arthritis.
\emph{Nature biotechnology} \textbf{31,} 142--7 (2013).

\hypertarget{ref-Yuan2014}{}
13. Yuan, W. \emph{et al.} An integrated epigenomic analysis for type 2
diabetes susceptibility loci in monozygotic twins. \emph{Nature
communications} \textbf{5,} 5719 (2014).

\hypertarget{ref-Tang2014}{}
14. Tang, Y. \emph{et al.} Genotype-based treatment of type 2 diabetes
with an \(\alpha\)2A-adrenergic receptor antagonist. \emph{Science
translational medicine} \textbf{6,} 257ra139 (2014).

\hypertarget{ref-Hong2015}{}
15. Hong, X. \emph{et al.} Genome-wide association study identifies
peanut allergy-specific loci and evidence of epigenetic mediation in US
children. \emph{Nature communications} \textbf{6,} 6304 (2015).

\hypertarget{ref-Claussnitzer2015}{}
16. Claussnitzer, M. \emph{et al.} FTO Obesity Variant Circuitry and
Adipocyte Browning in Humans. \emph{The New England journal of medicine}
\textbf{373,} 895--907 (2015).

\hypertarget{ref-Timpson2011}{}
17. Timpson, N. J. \emph{et al.} C-reactive protein levels and body mass
index: elucidating direction of causation through reciprocal Mendelian
randomization. \emph{International journal of obesity (2005)}
\textbf{35,} 300--8 (2011).

\hypertarget{ref-Shakhbazov2016}{}
18. Shakhbazov, K. \emph{et al.} Shared genetic control of expression
and methylation in peripheral blood. \emph{BMC genomics} \textbf{17,}
278 (2016).

\hypertarget{ref-Pierce2013}{}
19. Pierce, B. L. \& Burgess, S. Efficient design for Mendelian
randomization studies: subsample and 2-sample instrumental variable
estimators. \emph{American journal of epidemiology} \textbf{178,}
1177--84 (2013).

\hypertarget{ref-Ashenfelter1994}{}
20. Ashenfelter, O. \& Krueger, A. B. Estimates of the Economic Return
to Schooling from a New Sample of Twins. \emph{The American Economic
Review} \textbf{84,} 1157--1173 (1994).

\hypertarget{ref-LeCessie2012}{}
21. Cessie, S. le, Debeij, J., Rosendaal, F. R., Cannegieter, S. C. \&
Vandenbroucke, J. P. Quantification of bias in direct effects estimates
due to different types of measurement error in the mediator.
\emph{Epidemiology (Cambridge, Mass.)} \textbf{23,} 551--60 (2012).

\hypertarget{ref-Nagarajan2013}{}
22. Nagarajan, R. \& Scutari, M. Impact of noise on molecular network
inference. \emph{PloS one} \textbf{8,} e80735 (2013).

\hypertarget{ref-Shpitser2010}{}
23. Shpitser, I., VanderWeele, T. \& Robins, J. On the validity of
covariate adjustment for estimating causal effects. \emph{Proceedings of
the Twenty Sixth Conference on Uncertainty in Artificial Intelligence
(UAI-10)} 527--536 (2010).

\hypertarget{ref-Blakely2013}{}
24. Blakely, T., McKenzie, S. \& Carter, K. Misclassification of the
mediator matters when estimating indirect effects. \emph{Journal of
epidemiology and community health} \textbf{67,} 458--66 (2013).

\hypertarget{ref-Wang2015}{}
25. Wang, L. \& Michoel, T. Detection of regulator genes and eQTLs in
gene networks. \emph{arXiv} \textbf{arXiv:1512,} (2015).

\hypertarget{ref-Lagani2015}{}
26. Lagani, V., Triantafillou, S., Ball, G., Tegner, J. \& Tsamardinos,
I. in \emph{Uncertainty in biology: A computational modeling approach}
47 (Springer, 2015). at
\textless{}\href{https://books.google.com/books?id=8SLUCgAAQBAJ\%7B/\&\%7Dpgis=1}{https://books.google.com/books?id=8SLUCgAAQBAJ\{\textbackslash{}\&\}pgis=1}\textgreater{}

\hypertarget{ref-Houle2011}{}
27. Houle, D., Pélabon, C., Wagner, G. \& Hansen, T. Measurement and
meaning in biology. \emph{The Quarterly Review of Biology} \textbf{86,}
3--34 (2011).

\hypertarget{ref-Hernan2009}{}
28. Hernán, M. a \& Cole, S. R. Invited Commentary: Causal diagrams and
measurement bias. \emph{American journal of epidemiology} \textbf{170,}
959--62; discussion 963--4 (2009).

\hypertarget{ref-Harper2013}{}
29. Harper, K. N., Peters, B. a \& Gamble, M. V. Batch effects and
pathway analysis: two potential perils in cancer studies involving DNA
methylation array analysis. \emph{Cancer epidemiology, biomarkers \&
prevention : a publication of the American Association for Cancer
Research, cosponsored by the American Society of Preventive Oncology}
\textbf{22,} 1052--60 (2013).

\hypertarget{ref-Chen2013a}{}
30. Chen, Y.-a. \emph{et al.} Discovery of cross-reactive probes and
polymorphic CpGs in the Illumina Infinium HumanMethylation450
microarray. \emph{Epigenetics : official journal of the DNA Methylation
Society} \textbf{8,} 203--9 (2013).

\hypertarget{ref-Houseman2012}{}
31. Houseman, E. A. \emph{et al.} DNA methylation arrays as surrogate
measures of cell mixture distribution. \emph{BMC bioinformatics}
\textbf{13,} 86 (2012).

\hypertarget{ref-Ahima2013}{}
32. Ahima, R. S. \& Lazar, M. A. Physiology. The health risk of
obesity--better metrics imperative. \emph{Science (New York, N.Y.)}
\textbf{341,} 856--8 (2013).

\hypertarget{ref-RCoreTeam2015}{}
33. R Core Team. \emph{R: A Language and Environment for Statistical
Computing}. (R Foundation for Statistical Computing, 2015). at
\textless{}\url{https://www.r-project.org/}\textgreater{}

\hypertarget{ref-Henningsen2007}{}
34. Henningsen, A. \& Hamann, J. D. systemfit : A Package for Estimating
Systems of Simultaneous Equations in R. \emph{Journal of Statistical
Software} \textbf{23,} 1--40 (2007).

\hypertarget{ref-Steiger1980}{}
35. Steiger, J. H. Tests for comparing elements of a correlation matrix.
\emph{Psychological Bulletin} \textbf{87,} 245--251 (1980).

\hypertarget{ref-Revelle2015}{}
36. Revelle, W. \emph{psych: Procedures for Psychological, Psychometric,
and Personality Research}. (Northwestern University, 2015). at
\textless{}\url{http://cran.r-project.org/package=psych}\textgreater{}

\hypertarget{ref-Relton2012}{}
37. Relton, C. L. \& Davey Smith, G. Two-step epigenetic Mendelian
randomization: a strategy for establishing the causal role of epigenetic
processes in pathways to disease. \emph{International journal of
epidemiology} \textbf{41,} 161--76 (2012).

\hypertarget{ref-Burgess2014a}{}
38. Burgess, S., Freitag, D. F., Khan, H., Gorman, D. N. \& Thompson, S.
G. Using multivariable Mendelian randomization to disentangle the causal
effects of lipid fractions. \emph{PloS one} \textbf{9,} e108891 (2014).

\hypertarget{ref-Burgess2015}{}
39. Burgess, S., Daniel, R. M., Butterworth, A. S. \& Thompson, S. G.
Network Mendelian randomization: using genetic variants as instrumental
variables to investigate mediation in causal pathways.
\emph{International journal of epidemiology} \textbf{44,} 484--95
(2015).

\hypertarget{ref-Varbo2015}{}
40. Varbo, A. \emph{et al.} Remnant cholesterol, low-density lipoprotein
cholesterol, and blood pressure as mediators from obesity to ischemic
heart disease. \emph{Circulation research} \textbf{116,} 665--73 (2015).

\hypertarget{ref-Bowden2015}{}
41. Bowden, J., Davey Smith, G. \& Burgess, S. Mendelian randomization
with invalid instruments: effect estimation and bias detection through
Egger regression. \emph{International Journal of Epidemiology}
\textbf{In press,} (2015).
